### 2019-08-26 ch3_dynbenchmark/content.tex
refactorise project

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{dynbenchmark: A comparison of single-cell trajectory inference methods} 
\chaptermark{dynbenchmark: A comparison of single-cell trajectory inference methods.} % short title
\label{chap:dynbenchmark}

\begin{quote}
	\textbf{Abstract:} Trajectory inference approaches analyze genome-wide omics data from thousands of single cells and computationally infer the order of these cells along developmental trajectories. Although more than 70 trajectory inference tools have already been developed, it is challenging to compare their performance because the input they require and output models they produce vary substantially. Here, we benchmark 45 of these methods on 110 real and 229 synthetic datasets for cellular ordering, topology, scalability and usability. Our results highlight the complementarity of existing tools, and that the choice of method should depend mostly on the dataset dimensions and trajectory topology. Based on these results, we develop a set of guidelines to help users select the best method for their dataset. Our freely available data and evaluation pipeline (\href{https://benchmark.dynverse.org}{benchmark.dynverse.org}) will aid in the development of improved tools designed to analyze increasingly large and complex single-cell datasets.
\end{quote}

\vfill

Adapted from:\\
Saelens, W.$^*$, \textbf{Cannoodt, R.}$^*$, Todorov, H., and Saeys, Y. A comparison of single-cell trajectory inference methods. \textit{Nature Biotechnology} 37, 5 (2019), 547--554. \doi{10.1038/s41587-019-0071-9}.\\
{\footnotesize $^*$ Equal contribution}
\newpage

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/figure_1.pdf}
	\caption{
		\textbf{Overview of several key aspects of the evaluation.}
		\textbf{a}, A schematic overview of our evaluation pipeline. \textbf{b}, To make the trajectories comparable to each other, a common trajectory model was used to represent reference trajectories from the real and synthetic datasets, as well as any predictions of TI methods. \textbf{c}, Trajectories are automatically classified into one of seven trajectory types, with increasing complexity. \textbf{d}, We defined four metrics, each assessing the quality of a different aspect of the trajectory. The HIM score assesses the similarity between the two topologies, taking into account differences in edge lengths and degree distributions. The F1branches assesses the similarity of the assignment of cells onto branches. The cordist quantifies the similarity in cellular positions between two trajectories, by calculating the correlation between pairwise geodesic distances. Finally, wcorfeatures quantifies the agreement between trajectory differentially expressed features from the known trajectory and the predicted trajectory.
	}
	\label{fig:figure_1}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/supfigure_1.pdf}
	\caption{
		\textbf{A common interface for TI methods.}
		\textbf{a} The input and output of each TI method is standardized. As input, each TI method receives either raw or normalized counts, several parameters, and a selection of prior information. After its execution, a method uses one of the seven wrapper functions to transform its output to the common trajectory model. This common model then allows to perform common analysis functions on trajectory models produced by any TI method. \textbf{b} Illustrations of the specific transformations performed by each of the wrapper functions.
	}
	\label{fig:supfigure_1}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/figure_2.pdf}
	\caption{
		\textbf{A characterization of the 45 methods evaluated in this study and their overall evaluation results.}
		\textbf{a}, We characterized the methods according to the wrapper type, their required priors, whether the inferred topology is constrained by the algorithm (fixed) or a parameter (param), and the types of inferable topologies. The methods are grouped vertically based on the most complex trajectory type they can infer. \textbf{b}, The overall results of the evaluation on four criteria: accuracy using a reference trajectory on real and synthetic data, scalability with increasing number of cells and features, stability across dataset subsamples and quality of the implementation. Methods that errored on more than 50$\%$ of the datasets are not included in this figure and are shown instead in Supplementary Fig. 2.
	}
	\label{fig:figure_2}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/figure_3.pdf}
	\caption{	
		\textbf{Detailed results of the four main evaluation criteria: accuracy, scalability, stability and usability.}
		\textbf{a,} The names of the methods, ordered as in Figure \ref{fig:figure_2}. \textbf{b}, Accuracy of trajectory inference methods across metrics, dataset sources and dataset trajectory types. The performance of a method is generally more stable across dataset sources, but very variable depending on the metric and trajectory type. \textbf{c}, Predicted execution times for varying numbers of cells and features (no. of cells $\times$ no. of features). Predictions were made by training a regression model after running each method on bootstrapped datasets with varying numbers of cells and features. k, thousands; m, millions; cor, correlation. \textbf{d}, Stability results by calculating the average pairwise similarity between models inferred across multiple runs of the same method. \textbf{e}, Usability scores of the tool and corresponding manuscript, grouped per category. Off-the-shelf methods were directly implemented in R and thus do not have a usability score. 
	}
	\label{fig:figure_3}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/supfigure_3.pdf}
	\caption{
		\textbf{Accuracy of trajectory inference methods.}
		\textbf{a} Overall score for all methods across 339 datasets, colored by the source of the datasets. Black line indicates the mean. \textbf{b} Similarity between the overall scores of all dataset sources, compared to real datasets with a gold standard, across all methods (n = 46, after filtering out methods that errored too frequently). Shown in the top left is the Pearson correlation. \textbf{c} Bias in the overall score towards trajectory types for all methods across 339 datasets. Black line indicates the mean. \textbf{d} Distributions of the difference in size between predicted and reference topologies. A positive difference means that the topology predicted by the method is more complex than the one in the reference.
	}
	\label{fig:supfigure_3}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/figure_4.pdf}
	\caption{
		\textbf{Complementarity between different trajectory inference methods.}
		\textbf{a}, We assessed the likelihood for different combinations of methods to lead to a ‘top model’ (defined as a model with an overall score of at least 95$\%$ of the best model) when applied to all datasets. \textbf{b}, The likelihood for different combinations of methods to lead to a ‘top model’ was assessed separately on different trajectory types. For this figure, we did not include any methods requiring a cell grouping or a time course as prior information.
	}
	\label{fig:figure_4}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.9\linewidth]{fig/dynbenchmark/supfigure_4.pdf}
	\caption{
		\textbf{Scalability of trajectory inference methods.}
		\textbf{a} Three examples of average observed running times across five datasets (left) and the predicted running time (right). \textbf{b} Overview of the scalability results of all methods, ordered by their average predicted running time from (a). We predicted execution times and memory usage for each method with increasing number of features or cells, and used these values to classify each method into sublinear, linear, quadratic and superquadratic based on the shape of the curve.
	}
	\label{fig:supfigure_4}
\end{figure}

@@

Altogether, the scalability analysis indicated that the dimensions of the data are an important factor in the choice of method, and that method development should pay more attention to maintaining reasonable running times and memory usage.

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/figure_5.pdf}
	\caption{
		\textbf{Practical guidelines for method users.}
		As the performance of a method mostly depends on the topology of the trajectory, the choice of TI method will be primarily influenced by the user’s existing knowledge about the expected topology in the data. We therefore devised a set of practical guidelines, which combines the method’s performance, user friendliness and the number of assumptions a user is willing to make about the topology of the trajectory. Methods to the right are ranked according to their performance on a particular (set of) trajectory type. Further to the right are shown the accuracy (+: scaled performance $\geq$ 0.9, $\pm$: >0.6), usability scores (+:$\geq$0.9, $\pm$ $\geq$0.6), estimated running times and required prior information. k, thousands; m, millions.
	}
	\label{fig:figure_5}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/figure_6.pdf}
	\caption{
		\textbf{Demonstration of how a common framework for TI methods facilitates broad applicability using some example datasets.}
		Trajectories inferred by each method were projected to a common dimensionality reduction using multidimensional scaling. For each dataset, we also calculated a ‘consensus’ prediction, by calculating the cordist between each pair of models and picking the model with the highest score on average. \textbf{a}, The top methods applied on a dataset containing a linear trajectory of differentiation dendritic cells, going from MDP, CDP to PreDC. \textbf{b}, The top methods applied on a dataset containing a bifurcating trajectory of reprogrammed fibroblasts. \textbf{c}, A synthetic dataset generated by dyntoy, containing four disconnected trajectories. \textbf{d}, A synthetic dataset generated by dyngen, containing a cyclic trajectory.
	}
	\label{fig:figure_6}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_1.pdf}
	\caption{
		\textbf{An example trajectory that will be used throughout this section.}
		It contains contains four milestones (W to Z) and five cells (a to e).
	}
	\label{fig:snote1fig_1}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_2.pdf}
	\caption{
		\textbf{Showcase of three metrics to evaluate topologies: \textit{isomorphic}, \textit{edgeflip} and \textit{HIM}}
		\textbf{(a)} The used topologies. \textbf{(b)} The scores when comparing each pair of trajectory types. (c) Four datasets in which aan extra edge is added and made progressively longer. This shows how the HIM can take into account edge lengths.
	}
	\label{fig:snote1fig_2}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_3.pdf}
	\caption{
		\textbf{Mapping cells to their closest milestone or branch for the calculation of the {$\textit{F1}_{\textit{milestones}}$} and {$\textit{F1}_{\textit{branches}}$} .}
		To calculate the {$\textit{F1}_{\textit{milestones}}$}, cells are mapped towards the nearest milestone, i.e. the milestone with the highest milestone percentage. For the {$\textit{F1}_{\textit{branches}}$}, the cells are mapped to the closest edge.
	}
	\label{fig:snote1fig_3}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_4.pdf}
	\caption{
		\textbf{The calculation of geodesic distances on a small example trajectory.}
		\textbf{a)} A toy example containing four milestones (W to Z) and five cells (a to e). \textbf{b)} The corresponding milestone network, milestone percentages and regions of delayed commitment, when the toy trajectory is converted to the common trajectory model. \textbf{c)} The calculations made for calculating the pairwise geodesic distances. \textbf{d)} A heatmap representation of the pairwise geodesic distances.
	}
	\label{fig:snote1fig_4}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_5.pdf}
	\caption{
		\textbf{Determination of cell waypoints} \textbf{a)} Illustration of the stratified cell sampling using an example dataset (top). Each milestone, edge between two milestones and region of delayed commitment is seen as a collection of cells (middle), and the number of waypoints (100 in this case) are divided over each of these collection of cells (bottom). \textbf{b)} Accuracy versus time to calculate $\textit{cor}_{\textit{dist}}$. Shown are distributions over 100 random waypoint samples. The upper whisker of the boxplot extends from the hinge (75$\%$ percentile) to the largest value, no further than 1.5$\times$ the IQR of the hinge. The lower whisker extends from the hinge (25$\%$ percentile) to the smallest value, at most 1.5$\times$ the IQR of the hinge.
	}
	\label{fig:snote1fig_5}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_6.pdf}
	\caption{
		\textbf{Determination of cell waypoints.} 
		We generated different toy trajectory datasets with varying topologies and calculated the geodesic distances between all cells within the trajectory. We then used these distances as input for classical multidimensional scaling. This shows that the geodesic distances do not only contain information regarding the cell's positions, but also information on the lengths and wiring of the topology.
	}
	\label{fig:snote1fig_6}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_7.pdf}
	\caption{
		\textbf{The calculation of $\textit{NMSE}_{\textit{lm}}$ distances on a small example trajectory.} 
		The milestone percentages of the reference are predicted based on the milestone percentages of the prediction, using regression models such as linear regression or random forests. The predicted trajectory is then scored by comparing the mean-squared error (MSE) of this regression model with the baseline MSE where the prediction is the average milestone percentage.
	}
	\label{fig:snote1fig_7}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_8.pdf}
	\caption{
		\textbf{An illustration of ranking features based on their importance in a trajectory.} 
		(a) A MDS dimensionality reudction of a real dataset in which mouse embryonic fibroblasts (MEF) differentiate into Neurons and Myocytes. (b) The ranking of feature importances from high to low. The majority of features have a very low importance. (c) Some examples, which were also highlighted in b. Higher features in the ranking are clearly specific to certain parts of the trajectory, while features lower on the ranking have a more dispersed expression pattern.
	}
	\label{fig:snote1fig_8}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_9.pdf}
	\caption{
		\textbf{Effect of the number of trees parameter on the accuracy and variability of the $\textit{cor}_{\textit{features}}$.} 
		We used the dataset from Figure \ref{fig:snote1fig_8} and calculated the $\textit{cor}_{\textit{features}}$ after shuffling a percentage of cells.
	}
	\label{fig:snote1fig_9}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_10.pdf}
	\caption{
		\textbf{Effect of weighting the features based on their feature importance in the reference.} 
		We used the same dataset as in Figure \ref{fig:snote1fig_8}, and calculated the $\textit{cor}_{\textit{features}}$ after shuffling a percentage of cells.
	}
	\label{fig:snote1fig_10}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/dynbenchmark/snote1fig_11.pdf}
	\caption{
		\textbf{An illustration of how the difficulty of a dataset can influence the overall ranking.} 
		A decent method, which consistently ranks high on an easy and difficult dataset, does not get a high score when averaging. On the other hand, a method which ranks high on the easy dataset, but very low on the difficult dataset does get a high score on average. After normalising the scores (right), this problem disappears.
	}
	\label{fig:snote1fig_11}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=0.8\linewidth]{fig/dynbenchmark/snote1fig_12.pdf}
	\caption{
		\textbf{An example of the normalisation procedure.} 
		Shown are some results of a benchmarking procedure, where every row contains the scores of a particular method (red shading) on a particular dataset (blue shading), with a trajectory type (green shading) and dataset source (orange shading).
	}
	\label{fig:snote1fig_12}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.8\linewidth]{fig/dynbenchmark/snote1fig_13.pdf}
	\caption{
		\textbf{An example of the aggregation procedure.} 
		In consecutive steps we aggregated across \textbf{(a)} different datasets with the same source and trajectory type, \textbf{(b)} different dataset sources with the same trajectory type (weighted for the correlation of the dataset source with the real gold dataset source) and \textbf{(c)} all trajectory types.
	}
	\label{fig:snote1fig_13}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.8\linewidth]{fig/dynbenchmark/snote1fig_14.pdf}
	\caption{
		\textbf{An example of the averaging procedure.} 
		For each method, we calculated the geometric mean between its normalised and aggregated scores
	}
	\label{fig:snote1fig_14}
\end{figure}

>>>>

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/figure_1.pdf}
	\caption{
		\textbf{Overview of several key aspects of the evaluation.}
		\textbf{a}, A schematic overview of our evaluation pipeline. \textbf{b}, To make the trajectories comparable to each other, a common trajectory model was used to represent reference trajectories from the real and synthetic datasets, as well as any predictions of TI methods. \textbf{c}, Trajectories are automatically classified into one of seven trajectory types, with increasing complexity. \textbf{d}, We defined four metrics, each assessing the quality of a different aspect of the trajectory. The HIM score assesses the similarity between the two topologies, taking into account differences in edge lengths and degree distributions. The F1branches assesses the similarity of the assignment of cells onto branches. The cordist quantifies the similarity in cellular positions between two trajectories, by calculating the correlation between pairwise geodesic distances. Finally, wcorfeatures quantifies the agreement between trajectory differentially expressed features from the known trajectory and the predicted trajectory.
	}
	\label{fig:figure_1}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/supfigure_1.pdf}
	\caption{
		\textbf{A common interface for TI methods.}
		\textbf{a} The input and output of each TI method is standardized. As input, each TI method receives either raw or normalized counts, several parameters, and a selection of prior information. After its execution, a method uses one of the seven wrapper functions to transform its output to the common trajectory model. This common model then allows to perform common analysis functions on trajectory models produced by any TI method. \textbf{b} Illustrations of the specific transformations performed by each of the wrapper functions.
	}
	\label{fig:supfigure_1}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/figure_2.pdf}
	\caption{
		\textbf{A characterization of the 45 methods evaluated in this study and their overall evaluation results.}
		\textbf{a}, We characterized the methods according to the wrapper type, their required priors, whether the inferred topology is constrained by the algorithm (fixed) or a parameter (param), and the types of inferable topologies. The methods are grouped vertically based on the most complex trajectory type they can infer. \textbf{b}, The overall results of the evaluation on four criteria: accuracy using a reference trajectory on real and synthetic data, scalability with increasing number of cells and features, stability across dataset subsamples and quality of the implementation. Methods that errored on more than 50$\%$ of the datasets are not included in this figure and are shown instead in Supplementary Fig. 2.
	}
	\label{fig:figure_2}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/figure_3.pdf}
	\caption{	
		\textbf{Detailed results of the four main evaluation criteria: accuracy, scalability, stability and usability.}
		\textbf{a,} The names of the methods, ordered as in Figure \ref{fig:figure_2}. \textbf{b}, Accuracy of trajectory inference methods across metrics, dataset sources and dataset trajectory types. The performance of a method is generally more stable across dataset sources, but very variable depending on the metric and trajectory type. \textbf{c}, Predicted execution times for varying numbers of cells and features (no. of cells $\times$ no. of features). Predictions were made by training a regression model after running each method on bootstrapped datasets with varying numbers of cells and features. k, thousands; m, millions; cor, correlation. \textbf{d}, Stability results by calculating the average pairwise similarity between models inferred across multiple runs of the same method. \textbf{e}, Usability scores of the tool and corresponding manuscript, grouped per category. Off-the-shelf methods were directly implemented in R and thus do not have a usability score. 
	}
	\label{fig:figure_3}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/supfigure_3.pdf}
	\caption{
		\textbf{Accuracy of trajectory inference methods.}
		\textbf{a} Overall score for all methods across 339 datasets, colored by the source of the datasets. Black line indicates the mean. \textbf{b} Similarity between the overall scores of all dataset sources, compared to real datasets with a gold standard, across all methods (n = 46, after filtering out methods that errored too frequently). Shown in the top left is the Pearson correlation. \textbf{c} Bias in the overall score towards trajectory types for all methods across 339 datasets. Black line indicates the mean. \textbf{d} Distributions of the difference in size between predicted and reference topologies. A positive difference means that the topology predicted by the method is more complex than the one in the reference.
	}
	\label{fig:supfigure_3}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/figure_4.pdf}
	\caption{
		\textbf{Complementarity between different trajectory inference methods.}
		\textbf{a}, We assessed the likelihood for different combinations of methods to lead to a ‘top model’ (defined as a model with an overall score of at least 95$\%$ of the best model) when applied to all datasets. \textbf{b}, The likelihood for different combinations of methods to lead to a ‘top model’ was assessed separately on different trajectory types. For this figure, we did not include any methods requiring a cell grouping or a time course as prior information.
	}
	\label{fig:figure_4}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.9\linewidth]{fig/supfigure_4.pdf}
	\caption{
		\textbf{Scalability of trajectory inference methods.}
		\textbf{a} Three examples of average observed running times across five datasets (left) and the predicted running time (right). \textbf{b} Overview of the scalability results of all methods, ordered by their average predicted running time from (a). We predicted execution times and memory usage for each method with increasing number of features or cells, and used these values to classify each method into sublinear, linear, quadratic and superquadratic based on the shape of the curve.
	}
	\label{fig:supfigure_4}
\end{figure}

@@

Altogether, the scalability analysis indicated that the dimensions of the data are an important factor in the choice of method, and that method development should pay more attention to maintaining reasonable running times and memory usage.
§
\subsection{Stability}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/figure_5.pdf}
	\caption{
		\textbf{Practical guidelines for method users.}
		As the performance of a method mostly depends on the topology of the trajectory, the choice of TI method will be primarily influenced by the user’s existing knowledge about the expected topology in the data. We therefore devised a set of practical guidelines, which combines the method’s performance, user friendliness and the number of assumptions a user is willing to make about the topology of the trajectory. Methods to the right are ranked according to their performance on a particular (set of) trajectory type. Further to the right are shown the accuracy (+: scaled performance $\geq$ 0.9, $\pm$: >0.6), usability scores (+:$\geq$0.9, $\pm$ $\geq$0.6), estimated running times and required prior information. k, thousands; m, millions.
	}
	\label{fig:figure_5}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/figure_6.pdf}
	\caption{
		\textbf{Demonstration of how a common framework for TI methods facilitates broad applicability using some example datasets.}
		Trajectories inferred by each method were projected to a common dimensionality reduction using multidimensional scaling. For each dataset, we also calculated a ‘consensus’ prediction, by calculating the cordist between each pair of models and picking the model with the highest score on average. \textbf{a}, The top methods applied on a dataset containing a linear trajectory of differentiation dendritic cells, going from MDP, CDP to PreDC. \textbf{b}, The top methods applied on a dataset containing a bifurcating trajectory of reprogrammed fibroblasts. \textbf{c}, A synthetic dataset generated by dyntoy, containing four disconnected trajectories. \textbf{d}, A synthetic dataset generated by dyngen, containing a cyclic trajectory.
	}
	\label{fig:figure_6}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_1.pdf}
	\caption{
		\textbf{An example trajectory that will be used throughout this section.}
		It contains contains four milestones (W to Z) and five cells (a to e).
	}
	\label{fig:snote1fig_1}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_2.pdf}
	\caption{
		\textbf{Showcase of three metrics to evaluate topologies: \textit{isomorphic}, \textit{edgeflip} and \textit{HIM}}
		\textbf{(a)} The used topologies. \textbf{(b)} The scores when comparing each pair of trajectory types. (c) Four datasets in which aan extra edge is added and made progressively longer. This shows how the HIM can take into account edge lengths.
	}
	\label{fig:snote1fig_2}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_3.pdf}
	\caption{
		\textbf{Mapping cells to their closest milestone or branch for the calculation of the {$\textit{F1}_{\textit{milestones}}$} and {$\textit{F1}_{\textit{branches}}$} .}
		To calculate the {$\textit{F1}_{\textit{milestones}}$}, cells are mapped towards the nearest milestone, i.e. the milestone with the highest milestone percentage. For the {$\textit{F1}_{\textit{branches}}$}, the cells are mapped to the closest edge.
	}
	\label{fig:snote1fig_3}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_4.pdf}
	\caption{
		\textbf{The calculation of geodesic distances on a small example trajectory.}
		\textbf{a)} A toy example containing four milestones (W to Z) and five cells (a to e). \textbf{b)} The corresponding milestone network, milestone percentages and regions of delayed commitment, when the toy trajectory is converted to the common trajectory model. \textbf{c)} The calculations made for calculating the pairwise geodesic distances. \textbf{d)} A heatmap representation of the pairwise geodesic distances.
	}
	\label{fig:snote1fig_4}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_5.pdf}
	\caption{
		\textbf{Determination of cell waypoints} \textbf{a)} Illustration of the stratified cell sampling using an example dataset (top). Each milestone, edge between two milestones and region of delayed commitment is seen as a collection of cells (middle), and the number of waypoints (100 in this case) are divided over each of these collection of cells (bottom). \textbf{b)} Accuracy versus time to calculate $\textit{cor}_{\textit{dist}}$. Shown are distributions over 100 random waypoint samples. The upper whisker of the boxplot extends from the hinge (75$\%$ percentile) to the largest value, no further than 1.5$\times$ the IQR of the hinge. The lower whisker extends from the hinge (25$\%$ percentile) to the smallest value, at most 1.5$\times$ the IQR of the hinge.
	}
	\label{fig:snote1fig_5}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_6.pdf}
	\caption{
		\textbf{Determination of cell waypoints.} 
		We generated different toy trajectory datasets with varying topologies and calculated the geodesic distances between all cells within the trajectory. We then used these distances as input for classical multidimensional scaling. This shows that the geodesic distances do not only contain information regarding the cell's positions, but also information on the lengths and wiring of the topology.
	}
	\label{fig:snote1fig_6}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_7.pdf}
	\caption{
		\textbf{The calculation of $\textit{NMSE}_{\textit{lm}}$ distances on a small example trajectory.} 
		The milestone percentages of the reference are predicted based on the milestone percentages of the prediction, using regression models such as linear regression or random forests. The predicted trajectory is then scored by comparing the mean-squared error (MSE) of this regression model with the baseline MSE where the prediction is the average milestone percentage.
	}
	\label{fig:snote1fig_7}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_8.pdf}
	\caption{
		\textbf{An illustration of ranking features based on their importance in a trajectory.} 
		(a) A MDS dimensionality reudction of a real dataset in which mouse embryonic fibroblasts (MEF) differentiate into Neurons and Myocytes. (b) The ranking of feature importances from high to low. The majority of features have a very low importance. (c) Some examples, which were also highlighted in b. Higher features in the ranking are clearly specific to certain parts of the trajectory, while features lower on the ranking have a more dispersed expression pattern.
	}
	\label{fig:snote1fig_8}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_9.pdf}
	\caption{
		\textbf{Effect of the number of trees parameter on the accuracy and variability of the $\textit{cor}_{\textit{features}}$.} 
		We used the dataset from Figure \ref{fig:snote1fig_8} and calculated the $\textit{cor}_{\textit{features}}$ after shuffling a percentage of cells.
	}
	\label{fig:snote1fig_9}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_10.pdf}
	\caption{
		\textbf{Effect of weighting the features based on their feature importance in the reference.} 
		We used the same dataset as in Figure \ref{fig:snote1fig_8}, and calculated the $\textit{cor}_{\textit{features}}$ after shuffling a percentage of cells.
	}
	\label{fig:snote1fig_10}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=\linewidth]{fig/snote1fig_11.pdf}
	\caption{
		\textbf{An illustration of how the difficulty of a dataset can influence the overall ranking.} 
		A decent method, which consistently ranks high on an easy and difficult dataset, does not get a high score when averaging. On the other hand, a method which ranks high on the easy dataset, but very low on the difficult dataset does get a high score on average. After normalising the scores (right), this problem disappears.
	}
	\label{fig:snote1fig_11}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=0.8\linewidth]{fig/snote1fig_12.pdf}
	\caption{
		\textbf{An example of the normalisation procedure.} 
		Shown are some results of a benchmarking procedure, where every row contains the scores of a particular method (red shading) on a particular dataset (blue shading), with a trajectory type (green shading) and dataset source (orange shading).
	}
	\label{fig:snote1fig_12}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.8\linewidth]{fig/snote1fig_13.pdf}
	\caption{
		\textbf{An example of the aggregation procedure.} 
		In consecutive steps we aggregated across \textbf{(a)} different datasets with the same source and trajectory type, \textbf{(b)} different dataset sources with the same trajectory type (weighted for the correlation of the dataset source with the real gold dataset source) and \textbf{(c)} all trajectory types.
	}
	\label{fig:snote1fig_13}
\end{figure}

@@

\begin{figure}
	\centering\includegraphics[width=0.8\linewidth]{fig/snote1fig_14.pdf}
	\caption{
		\textbf{An example of the averaging procedure.} 
		For each method, we calculated the geometric mean between its normalised and aggregated scores
	}
	\label{fig:snote1fig_14}
\end{figure}
